---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

The Seattle Public Library has provided us with a dataset of items that have been checked out at least 10 times a month from 2017-2023. This dataset is a comprehensive record of popular books, media, and other materials that have been in high demand at the library over the past few years. This information is valuable for understanding the reading and media consumption habits of the Seattle community, as well as for predicting future trends in library usage.

###### Description of Trends/Books/Items to Analyze:

**Genre Popularity**: This trend will analyze the most popular genres among the checked-out items. It will provide insight into what kind of literature is in demand and help the library to curate its collection accordingly.

**Yearly Trends**: This trend will analyze the change in item checkouts throughout the decades We can determine if there is any certain demand for specific genres, which can help the library in planning its collection development.

**Format Preferences**: This trend will analyze the preferred format of items checked out, such as eBooks, audiobooks, or physical books. Understanding the preferred format of items can help the library to make informed decisions about its technology and budget.

**Author Popularity**: This trend will analyze the most popular authors among the checked-out items. It can help the library to understand what authors are in demand and assist in selecting which author events to host.

**JK Rowling/The Harry Potter Series**: Although JK Rowling is a horrible person, Harry Potter is one of my favorite seies and I would love to analyze fluctuations in its popularity (generation or controversy might affect it) and perhaps compare it to Rowling's other releases.

### Summary Information

Based on the analysis of the Seattle Public Library checkouts dataset, it was found that books were the most checked out material type, still reigning supreme (ebooks have not taken over yet)! The total number of checkouts fluctuated, with 2020 being the year with the lowest number of checkouts, which is surprising since everyone was at home and presumably reading more. However, 2020 did have the most checkouts for the Harry Potter books specifically, proving that they are still remaining relevant. Ebook checkouts have been on the rise, with the least amount of ebook checkouts in 2017 and the most in 2020, which makes sense since people were sitting at home because of COVID and checking out online copies. 2020 was also the year with the most nonfiction checkouts. These findings are interesting because they suggest that 2020 was a busy year for the SPL, with high numbers of checkouts for various genres and materials. However, the total amount of checkouts was the lowest in the last 6 years.

### The Dataset

The Seattle Public Library (SPL) collected and published the data on their website. The dataset contains information on checkouts from 2015 to the present, including the number of checkouts, types of materials, and the checkout year. The data includes information on books, ebooks, audiobooks, and other materials. The data was generated by recording the number of checkouts for each material type each year. The SPL likely collected this data to better understand how their resources were being used by the community and to inform future collection development decisions.

When working with this data, there are ethical considerations to keep in mind. For example, the dataset contains information about the titles that individuals checked out, which could be considered private. Researchers must ensure that they are complying with any relevant data privacy laws and regulations, and they must take steps to protect individuals' privacy, yet the SPL already does that exceptionally well. It is also important to ensure that any conclusions drawn from the data are not discriminatory in nature, and that the data is being used for the benefit of the community.

Possible limitations or problems with this data include the fact that it only represents one library system in one city, so the findings may not be generalizable to other areas. Additionally, the dataset does not include demographic information, which could limit the insights that can be gained from the data. Finally, the dataset only includes information on physical and digital checkouts, and does not include information on in-person visits to the library or other activities that library users may engage in.

### First Trends Over Time Chart

This chart visually represents the trends in total checkouts per year, and compare the trend in ebook checkouts to the trend in all checkouts. From the chart, we can see that there is a general upward trend in total checkouts from 2017 to 2019, with a slight dip in 2020, which can be attributed to the COVID-19 pandemic. However, checkouts increased again in 2021 and 2022, which could be due to the gradual lifting of pandemic restrictions. We can also see that the trend in ebook checkouts is generally upward as well, although it is less pronounced than the trend in all checkouts. This suggests that while ebooks are becoming increasingly popular, physical books are still in high demand. Overall, this chart provides a clear and concise visualization of the trends in checkout patterns at the Seattle Public Library.

``````{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart1_example.R")}

```

### Second Trends Over Time Chart

I included this chart to visualize the popularity of each Harry Potter book over time. The chart shows the total number of checkouts for each book from 2017 to 2023, which is useful for libraries or booksellers who want to understand which books are still in demand.

One pattern that emerged from the chart is that the earlier books in the series (Sorcerer's Stone, Chamber of Secrets, and Prisoner of Azkaban) have remained popular over time, with consistently high checkouts each year. In contrast, the later books (Half-Blood Prince and Deathly Hallows) have lower checkouts, which may be due to a combination of factors such as reader fatigue with the series or competition from other popular books.

Another pattern that emerged is that the checkouts for each book tend to peak in the year of a major movie release. For example, the checkouts for Sorcerer's Stone and Chamber of Secrets spiked in 2020, the year that the Fantastic Beasts sequel was released, while the checkouts for Goblet of Fire spiked in 2021, the year that the 20th anniversary edition of the book was released.

``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Third Trends by Material Type

This bar chart gains insights into the top 6 material types that have the highest total checkouts from 2017-2023. It's specifically top 6 because 7 and onwards' numbers are too small and insignificant in relation. The chart shows that books are the most checked out material type, followed by ebooks, audiobooks, and DVDs. Magazines and music CDs have the lowest total checkouts among the top 6 materials.

The patterns that emerged from this chart are that books and ebooks are the most popular material types among library patrons, which could be due to their accessibility and convenience. It's also interesting to note that audiobooks have a high total checkout rate, which suggests that patrons may be utilizing them for their convenience and portability. DVDs, on the other hand, have a lower total checkout rate compared to books and ebooks, which could be attributed to the rise of online streaming services.

Overall, this chart provides valuable insights into the most popular material types among library patrons, which can help inform decisions on library collection development and resource allocation.

``````{r, echo = FALSE, code = readLines("chart3_example.R")}
```


